COFFEE: an Optimizing Compiler for Finite Element Local Assembly

### 2014-06-23 TACO-paper/coffee.tex
Clarify why we do not consider high order methods

We focus on improving the performance of local assembly (``assembly'', in the following), a fundamental step in the finite element method that covers an important fraction of the overall computation run-time, often in the range 30-60$\%$. During the assembly phase, the solution of the PDE is approximated by executing a suitable kernel over all elements in the discretized domain. A kernel's working set is usually small enough to fit the L1 cache; it might need L2 cache when high-order methods are employed. However, we do not consider the latter case. An assembly kernel is characterized by the presence of an affine, often non-perfect loop nest, in which individual loops are rather small: their trip count rarely exceeds 30, and may be as low as 3 for low order methods. In the innermost loop, a problem-specific expression evaluates a two dimensional array, representing the result of local assembly in an element of the discretized domain. With such a kernel structure, we focus on aspects like minimization of floating-point operations, register allocation and instruction-level parallelism, especially in the form of SIMD vectorization.

>>>>

We focus on improving the performance of local assembly (or ``assembly'', in the following), a fundamental step in the finite element method that covers an important fraction of the overall computation run-time, often in the range 30-60$\%$. During the assembly phase, the solution of the PDE is approximated by executing a suitable kernel over all elements in the discretized domain. In this work, we focus on relatively low order finite element methods, in which an assembly kernel's working set is usually small enough to fit the L1 cache. Low order methods are by no means exotic: they are employed in a wide variety of fields, including climate and ocean modeling, computational fluid dynamics, and structural mechanics. High order methods are often based upon special formulations of the finite element method, which result in the adoption of different sets of algorithms to solve PDEs~\cite{SPENCER}, and therefore are excluded from our study.

An assembly kernel is characterized by the presence of an affine, often non-perfect loop nest, in which individual loops are rather small: their trip count rarely exceeds 30, and may be as low as 3 for low order methods. In the innermost loop, a problem-specific expression evaluates a two dimensional array, representing the result of local assembly in an element of the discretized domain. With such a kernel structure, we focus on aspects like minimization of floating-point operations, register allocation and instruction-level parallelism, especially in the form of SIMD vectorization.
